Backscatter Modulation Design for Symbiotic Radio Networks

### 2021-11-29 main.tex
:pencil: Update energy distribution and formulate problem

			To explicitly specify each backscatter symbol combination, we label it by the corresponding modulation index set. For tag $k$ at state $m_k$, $\forall k \in \mathcal{K}$, we define the modulation index set as $m_{\mathcal{K}} \triangleq \{m_k : k \in \mathcal{K}\}$ and the tag input combination as $\bar{c}_{m_{\mathcal{K}}} \triangleq \{\bar{c}_{m_k} : k \in \mathcal{K}\}$. Since any $\bar{c}_{m_{\mathcal{K}}}$ remains constant per $N$ primary symbols, the received signal at block $n$ is only subject to the variation of the primary source $s[n]$ and AWGN $w[n]$, and thus follows CSCG distribution $y_{m_{\mathcal{K}}}[n] \sim \mathcal{CN}\left(0,\sigma_{m_{\mathcal{K}}}^2\right)$ with variance
			\begin{equation}
				\sigma_{m_{\mathcal{K}}}^2 = \Bigl\lvert \underbrace{\bigl(\boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H \bar{c}_{m_k}\bigr)}_{\boldsymbol{h}_{\mathrm{E}}^H(\bar{c}_{m_{\mathcal{K}}})} \boldsymbol{w} \Bigr\rvert^2 + \sigma_w^2,
				\label{eq:receive_variance}
			\end{equation}
			which denotes the expectation of the received power per primary block under tag modulation index set $m_{\mathcal{K}}$, and can be numerically evaluated based on the precoder, tag, channel, and noise knowledge. For the ease of exposition, we sort all possible $\sigma_{m_{\mathcal{K}}}^2$ in ascending order by a one-to-one mapping $m_{\mathcal{K}} \mapsto i$, denote the hypothesis that the tag input combination at status $i$ as $\mathcal{H}_i$, and define $\sigma_i^2 \le \sigma_j^2$ for $i < j$, $i,j \in \mathcal{M^K} \triangleq \{1,\ldots,M^K\}$.\footnote{When more than one modulation index sets yield the same energy level, the mapping is not unique and the detection fails to separate them. This blind spot issue can be mitigated by multi-antenna techniques.} For the optimal Maximum-Likelihood (ML) detector, the likelihood ratio between hypotheses $\mathcal{H}_i$ and $\mathcal{H}_j$ is \cite{Qian2019}
			\begin{equation}
				\Lambda_{i,j}(\boldsymbol{y}) = \frac{f(\boldsymbol{y} \mid \mathcal{H}_i)}{f(\boldsymbol{y} \mid \mathcal{H}_j)} = \left( \frac{\sigma_j^2}{\sigma_i^2} \right)^N \exp \left( \frac{\sigma_i^2 - \sigma_j^2}{\sigma_i^2 \sigma_j^2} z \right),
				\label{eq:likelihood_ratio}
			\end{equation}
			where $f(\boldsymbol{y} \mid \mathcal{H}_i)$ and $f(\boldsymbol{y} \mid \mathcal{H}_j)$ are respectively the conditional probability density function of receiving $\boldsymbol{y}$ under hypothesis $\mathcal{H}_i$ and $\mathcal{H}_j$, and $z \triangleq \lVert \boldsymbol{y} \rVert^2$ is the received signal energy over $N$ primary blocks. Obviously, the likelihood ratio only depends on $z$ and the energy detection is thus optimal. The corresponding decision rule is
			\begin{equation}
				\Lambda_{i,j}(z) \underset{\mathcal{H}_j}{\overset{\mathcal{H}_i}{\lessgtr}} 1 \iff z \underset{\mathcal{H}_j}{\overset{\mathcal{H}_i}{\lessgtr}} T_{i,j},
				\label{eq:decision_rule}
			\end{equation}
			where the detection threshold between $\mathcal{H}_i$ and $\mathcal{H}_j$ is
			\begin{equation}
				T_{i,j} \triangleq N \frac{\sigma_i^2 \sigma_j^2}{\sigma_i^2 - \sigma_j^2} \log \frac{\sigma_i^2}{\sigma_j^2}.
				\label{eq:detection_threshold}
			\end{equation}

			Accordingly, the decision region of hypothesis $\mathcal{H}_i$ is
			\begin{equation}
				\mathcal{R}_i \triangleq [T_{i-1, i}, T_{i, i+1}),
				\label{eq:decision_region}
			\end{equation}
			and we manually define $T_{0,1} \triangleq 0$ and $T_{M^K,M^K+1} \triangleq \infty$. Let $f(z \mid \mathcal{H}_i)$ be the conditional probability density function of receiving $z$ under hypothesis $\mathcal{H}_i$. Since $z$ follows chi-squared distribution with $2N$ degrees of freedom, it holds that
			\begin{equation}
				f(z \mid \mathcal{H}_i) = \frac{z^{N-1} e^{-z/\sigma_i^2}}{\sigma_i^{2N} \Gamma(N)},
				\label{eq:energy_distribution}
			\end{equation}
			where $\Gamma(\cdot)$ is the gamma function. As such, we can formulate an equivalent point-to-point discrete memoryless channel from tag input combination alphabet $\mathcal{M^K}$ to received energy level alphabet $\mathcal{M^K}$ as a transition matrix, where the probability of observing energy level $j$ under tag input combination $i$ is\footnote{For simplicity, we assume there exists at least one feasible precoder that produces distinct received energy levels for all tag input combinations. [TODO] Is this necessary? Maybe we can keep those same energy levels and avoid them by input distribution design.}
			\begin{equation}
				P(\bar{z}_j \mid \bar{c}_i) = P(z \in \mathcal{R}_j \mid \mathcal{H}_i) = \int_{\mathcal{R}_j} f(z \mid \mathcal{H}_i) \dd z.
				\label{eq:point_to_point_channel}
			\end{equation}

@@

			Evidently, for a given discrete memoryless MAC with transition probability $P(z \mid c_{\mathcal{K}})$, the information function $I(\bar{c}_{m_{\mathcal{K}}} ; z)$ and mutual information $I(c_{\mathcal{K}} ; z)$ only depend on the tag input distribution $P_k(\bar{c}_{m_k})$, $\forall m_k \in \mathcal{M}, k \in \mathcal{K}$. Therefore, we can obtain the backscatter sum-capacity by maximizing the mutual information over the tag input distributions as
			\begin{equation}
				R_c = \max_{P(c_{\mathcal{K}})} I(c_{\mathcal{K}} ; z) \ [\si{bpcu}].
			\end{equation}

			% On top of them, the mutual information and marginal mutual information can be written respectively as\footnote{Please be aware that $c_{\mathcal{K}}$, $c_{\mathcal{S}}$, $z$ are random variables, while $\bar{c}_{m_{\mathcal{K}}}$ (and $\bar{c}_i$), $\bar{c}_{m_{\mathcal{S}}}$, $\bar{z}_j$ represent the corresponding instances.}
			% \begin{align}
			% 	I(c_{\mathcal{K}} ; z)
			% 	& = \mathbb{E}_{c_{\mathcal{S}}} \left[I_{\mathcal{S}}(\bar{c}_{m_{\mathcal{S}}} ; z)\right],\\
			% 	I(c_{\mathcal{S}} ; z)
			% 	& = \mathbb{E}_{c_{\mathcal{S}}} \left[I(\bar{c}_{m_{\mathcal{S}}} ; z)\right].
			% \end{align}

			% On top of \eqref{eq:information_function} and \eqref{eq:marginal_information_function}, the mutual information can be expressed as\footnote{Please be aware that $c_{\mathcal{K}}$, $c_{\mathcal{S}}$, $z$ are random variables, while $\bar{c}_{m_{\mathcal{K}}}$ (and $\bar{c}_i$), $\bar{c}_{m_{\mathcal{S}}}$, $\bar{z}_j$ represent the corresponding instances.}
			% \begin{equation}
			% 	I(c_{\mathcal{K}} ; z) = \mathbb{E}_{c_{\mathcal{K}}} \left[I(\bar{c}_{m_{\mathcal{K}}} ; z)\right] = \mathbb{E}_{c_{\mathcal{S}}} \left[I_{\mathcal{S}}(\bar{c}_{m_{\mathcal{S}}} ; z)\right].
			% \end{equation}

			% \begin{equation}
			% 	I_{\mathcal{S}}(\bar{c}_{m_{\mathcal{S}}} ; z) = \sum_{\mathcal{S}^c} P_{\mathcal{S}^c}(\bar{c}_{m_{\mathcal{S}^c}}) I(\bar{c}_{m_{\mathcal{K}}} ; z),
			% \end{equation}
			% where $\mathcal{S}^c \triangleq \mathcal{K} \setminus \mathcal{S}$.

			% and the mutual information can be expressed as
			% \begin{equation}
			% 	I(\boldsymbol{c} ; z) = \mathbb{E}_{\boldsymbol{c}}
			% \end{equation}

			% function that evaluates the information associated with individual

			% For $\mathcal{S} \subset \mathcal{K}$, denote the input probability distribution that user $r_{\mathcal{S}}(\bar{c}_{m_{\mathcal{S}}})$
		\end{subsection}

		\begin{subsection}{Primary ergodic capacity}
			Once the tag symbol set $\bar{c}_{m_{\mathcal{K}}}$ is successfully detected, the user can eliminate the backscatter uncertainty and obtain the equivalent primary channel by \eqref{eq:equivalent_primary_channel}. From the perspective of primary transmission, the tags adjust the propagation environment in a potentially beneficial manner, and create artificial channel variation within each fading block. As such, the ergodic primary capacity can be expressed as
			\begin{align}
				R_s
				& = \mathbb{E}_{c_{\mathcal{K}}} \left[\log_2 \left(1 + \frac{\lvert \boldsymbol{h}_{\mathrm{E}}^H(c_{\mathcal{K}}) \boldsymbol{w} \rvert^2}{\sigma_w^2}\right)\right]\nonumber\\
				& = \sum_{m_{\mathcal{K}}} \prod_{k \in \mathcal{K}} P_k(\bar{c}_{m_k}) \log_2 \left(1 + \frac{\lvert \boldsymbol{h}_{\mathrm{E}}^H(\bar{c}_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2}{\sigma_w^2}\right) \ [\si{bps/Hz}].
				\label{eq:primary_rate}
			\end{align}
			which is also a function of the tag input distribution $P_k(\bar{c}_{m_k})$, $\forall m_k \in \mathcal{M}, k \in \mathcal{K}$.
		\end{subsection}
		\label{se:system_model}
	\end{section}

	\begin{section}{Tag Input and Precoder Design}
		In this section, we exploit tag input and precoder design to characterize the achievable region of the primary rate and the backscatter sum-rate.
		\begin{subsection}{Backscatter sum-rate}

		\end{subsection}
	\end{section}

>>>>

			To explicitly specify each backscatter symbol combination, we label it by the corresponding modulation index set. For tag $k$ at state $m_k \in \mathcal{M}$, $\forall k \in \mathcal{K}$, we define the modulation index set as $m_{\mathcal{K}} \triangleq \{m_k : k \in \mathcal{K}\}$ and the tag input combination as $\bar{c}_{m_{\mathcal{K}}} \triangleq \{\bar{c}_{m_k} : k \in \mathcal{K}\}$. Since any $\bar{c}_{m_{\mathcal{K}}}$ remains constant per $N$ primary symbols, the received signal at block $n$ is only subject to the variation of the primary source $s[n]$ and AWGN $w[n]$, and thus follows CSCG distribution $y_{m_{\mathcal{K}}}[n] \sim \mathcal{CN}\left(0,\sigma_{m_{\mathcal{K}}}^2\right)$ with variance
			\begin{equation}
				\sigma_{m_{\mathcal{K}}}^2 = \Bigl\lvert \underbrace{\bigl(\boldsymbol{h}_{\mathrm{D}}^H + \sum_{k \in \mathcal{K}} \sqrt{\alpha_k} \boldsymbol{h}_{\mathrm{C},k}^H \bar{c}_{m_k}\bigr)}_{\boldsymbol{h}_{\mathrm{E}}^H(\bar{c}_{m_{\mathcal{K}}})} \boldsymbol{w} \Bigr\rvert^2 + \sigma_w^2,
				\label{eq:receive_variance}
			\end{equation}
			which denotes the expectation of the received power per primary block under tag modulation index set $m_{\mathcal{K}}$, and can be numerically evaluated based on the precoder, tag, channel, and noise knowledge. For the ease of exposition, we sort all possible $\sigma_{m_{\mathcal{K}}}^2$ in ascending order by a one-to-one mapping $m_{\mathcal{K}} \mapsto i$, denote the hypothesis that the tag input combination at status $i$ as $\mathcal{H}_i$, and define $\sigma_i^2 \le \sigma_j^2$ for $i < j$, $i,j \in \mathcal{M^K} \triangleq \{1,\ldots,M^K\}$.\footnote{When more than one modulation index sets yield the same energy level, the mapping is not unique and the detection fails to separate them. This blind spot issue can be mitigated by multi-antenna techniques.} For the optimal Maximum-Likelihood (ML) detector, the likelihood ratio between hypotheses $\mathcal{H}_i$ and $\mathcal{H}_j$ is \cite{Qian2019}
			\begin{equation}
				\Lambda_{i,j}(\boldsymbol{y}) = \frac{f(\boldsymbol{y} \mid \mathcal{H}_i)}{f(\boldsymbol{y} \mid \mathcal{H}_j)} = \left( \frac{\sigma_j^2}{\sigma_i^2} \right)^N \exp \left( \frac{\sigma_i^2 - \sigma_j^2}{\sigma_i^2 \sigma_j^2} z \right),
				\label{eq:likelihood_ratio}
			\end{equation}
			where $f(\boldsymbol{y} \mid \mathcal{H}_i)$ and $f(\boldsymbol{y} \mid \mathcal{H}_j)$ are respectively the conditional probability density function of receiving $\boldsymbol{y}$ under hypothesis $\mathcal{H}_i$ and $\mathcal{H}_j$, and $z \triangleq \lVert \boldsymbol{y} \rVert^2$ is the received signal energy over $N$ primary blocks. Obviously, the likelihood ratio only depends on $z$ and the energy detection is thus optimal. The corresponding decision rule is
			\begin{equation}
				\Lambda_{i,j}(z) \underset{\mathcal{H}_j}{\overset{\mathcal{H}_i}{\lessgtr}} 1 \iff z \underset{\mathcal{H}_j}{\overset{\mathcal{H}_i}{\lessgtr}} T_{i,j},
				\label{eq:decision_rule}
			\end{equation}
			where the detection threshold between $\mathcal{H}_i$ and $\mathcal{H}_j$ is
			\begin{equation}
				T_{i,j} \triangleq N \frac{\sigma_i^2 \sigma_j^2}{\sigma_i^2 - \sigma_j^2} \log \frac{\sigma_i^2}{\sigma_j^2}.
				\label{eq:detection_threshold}
			\end{equation}

			Accordingly, the decision region of hypothesis $\mathcal{H}_i$ is
			\begin{equation}
				\mathcal{R}_i \triangleq [T_{i-1,i}, T_{i,i+1}),
				\label{eq:decision_region}
			\end{equation}
			and we manually define $T_{0,1} \triangleq 0$ and $T_{M^K,M^K+1} \triangleq \infty$. Let $f(z \mid \mathcal{H}_i)$ be the conditional probability density function of receiving $z$ under hypothesis $\mathcal{H}_i$. Correspondingly, $z$ is the sum of $N$ i.i.d. exponential variables each with mean $\sigma_i^2$, and thus follows Erlang distribution with shape $N$ and scale $\sigma_i^2$
			\begin{equation}
				f(z \mid \mathcal{H}_i) = \frac{z^{N-1} e^{-z/\sigma_i^2}}{\sigma_i^{2N} (N-1)!}.
				\label{eq:energy_distribution}
			\end{equation}

			On top of \eqref{eq:decision_region} and \eqref{eq:energy_distribution}, we can formulate an equivalent point-to-point discrete memoryless channel from tag input combination alphabet $\mathcal{M^K}$ to received energy level alphabet $\mathcal{M^K}$ as a transition matrix, where the probability of observing energy level $j$ under tag input combination $i$ is\footnote{For simplicity, we assume there exists at least one feasible precoder that produces distinct received energy levels for all tag input combinations. [TODO] Is this necessary? Maybe we can keep those same energy levels and avoid them by input distribution design.}
			\begin{equation}
				P(\bar{z}_j \mid \bar{c}_i) = P(z \in \mathcal{R}_j \mid \mathcal{H}_i) = \int_{\mathcal{R}_j} f(z \mid \mathcal{H}_i) \dd z.
				\label{eq:point_to_point_channel}
			\end{equation}

@@

			Evidently, for a given discrete memoryless MAC with transition probability $P(z \mid c_{\mathcal{K}})$, the information function $I(\bar{c}_{m_{\mathcal{K}}} ; z)$ and mutual information $I(c_{\mathcal{K}} ; z)$ only depend on the tag input distribution $P_k(\bar{c}_{m_k})$, $\forall m_k \in \mathcal{M}$, $\forall k \in \mathcal{K}$. Therefore, we can obtain the backscatter sum-capacity by maximizing the mutual information over the tag input distributions as
			\begin{equation}
				R_c = \max_{P(c_{\mathcal{K}})} I(c_{\mathcal{K}} ; z) \ [\si{bpcu}].
			\end{equation}
		\end{subsection}

		\begin{subsection}{Primary ergodic capacity}
			Once the tag symbol set $\bar{c}_{m_{\mathcal{K}}}$ is successfully decoded, the user can eliminate the backscatter uncertainty and model it within the equivalent primary channel by \eqref{eq:equivalent_primary_channel}. From the perspective of primary transmission, the tags adjust the propagation environment in a potentially beneficial manner, and create artificial channel variation within each fading block. As such, the ergodic primary capacity can be expressed as
			\begin{align}
				R_s
				& = \mathbb{E}_{c_{\mathcal{K}}} \left[\log_2 \left(1 + \frac{\lvert \boldsymbol{h}_{\mathrm{E}}^H(c_{\mathcal{K}}) \boldsymbol{w} \rvert^2}{\sigma_w^2}\right)\right]\nonumber\\
				& = \sum_{m_{\mathcal{K}}} \prod_{k \in \mathcal{K}} P_k(\bar{c}_{m_k}) \log_2 \left(1 + \frac{\lvert \boldsymbol{h}_{\mathrm{E}}^H(\bar{c}_{m_{\mathcal{K}}}) \boldsymbol{w} \rvert^2}{\sigma_w^2}\right) \ [\si{bps/Hz}].
				\label{eq:primary_rate}
			\end{align}
			which is also a function of the tag input distribution $P(c_{\mathcal{K}})$.
		\end{subsection}
	\end{section}

	\begin{section}{Tag Input and Precoder Design}
		\begin{subsection}{Primary ergodic-secondary sum rate region}
			To achieve a flexible tradeoff between the primary and backscatter links, we aim to optimize the precoder and tag input distribution to maximize the weighted sum of the ergodic primary rate and the total backscatter rate, under the average transmit power and probability constraints
			\begin{maxi!}
				{\scriptstyle{{P(c_{\mathcal{K}})},{\boldsymbol{w}}}}{\rho R_s + (1 - \rho) R_c}{\label{op:rate_region}}{\label{ob:weighted_sum_rate}}
				\addConstraint{\lVert \boldsymbol{w} \rVert^2}{\le P\label{co:average_transmit_power}}
				\addConstraint{\sum_{m_k} P_k(\bar{c}_{m_k})}{=1,\quad\label{co:sum_probability}}{\forall k \in \mathcal{K}}
				\addConstraint{P_k(\bar{c}_{m_k})}{\ge 0,\quad\label{co:nonnegative_probability}}{\forall m_k \in \mathcal{M}, \ \forall k \in \mathcal{K}}
			\end{maxi!}
			where $\rho \in [0,1]$ denotes the priority of the primary link. As problem \eqref{op:rate_region} is not jointly convex over $\boldsymbol{w}$ and $P(c_{\mathcal{K}})$, we propose an alternating optimization (AO)-based algorithm that iteratively retrieve the global optimal precoder and capacity-achieving input distribution in closed forms.
		\end{subsection}

		\begin{subsection}{Precoder}
			TODO
		\end{subsection}
	\end{section}
